Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint

### 2014-11-14 egpaper_final.tex
uncomment FFT

\subsection{Overview}
Our method attempts to match CAD models

Our method is in line with \cite{Aubry14, Lim14} in the sense that we
use only renderings, not real images. However, rather than using fixed
sized mid-level patches, we use whole objects and make root
templates. By using root templates, there is no need to learn a
discriminative patch dictionary or to calibrate the patches and an
additional inference stage is not required. However, whitening the
whole template is prohibitive for several reasons. All templates have
different aspect ratios making it difficult to generate the covariance
matrix $\Sigma$. Since the root templates are larger than mid-level
patches, a very large auto-covariance $\Gamma$ is required. Lastly,
time complexity of decorrelating a template with $n$ cells increases
as $O(n^3)$, resulting in slow template generation and large residuals
due to numerical errors accumulated while decomposing the covariance
matrix $\Sigma$.

To overcome these difficulties, we propose the Non-Zero Whitened
Histograms of Orientations (NZ-WHO) with iterative Conjugate Gradient to
generate templates in real time and enrich rectangular boxes with high
quality 2D-3D matching. The 2D-3D matching provides a CAD model,
sub-category, rendering, depth, viewpoint and focal length.

Our method can work as a standalone 2D-3D matching algorithm, or even
as a standalone ensemble of exemplar-based object detectors (Figure
\ref{fig:front}) but its real power comes from combining generic
object detectors and complement rectangular bounding boxes. Our method
is simple in design; is several orders of magnitude faster than
traditional method; requires only CAD models; does not need neither images nor
calibration, yet is powerful in giving high-quality metadata. These
properties make our method ideal for complementing object detections.

To complement detection bounding boxes, we first search possible 2D-3D
matching by running cached discrete viewpoint templates. Then, for the
most probable match, we refine the initial matching using Metropolis
Hastings. Since it requires iterations of proposal and validation, we
present real time analysis of NZ-WHO template generation to show that
this can be done on-the-fly. The fine tuning stage process jointly
optimizes translation, yaw, roll, pitch, scale, and focal length
continuously and also searches for different CAD models.

We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), that are both more discriminative than WHO, and can be generated in 70ms (compared to several seconds for WHO). This speed up means we can generate templates on the fly, allowing our approach to evaluate arbitrary viewpoint hypotheses. 
\scream{chris:We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), an improvement over WHO which is both more discriminative than WHO and can be computed many orders of magnitude faster than WHO (70ms as opposed to to several seconds for WHO). This speedup allows us to generate NZ-WHO templates on the fly, allowing us to quickly evaluate arbitrary viewpoint hypotheses.}

The computational efficiency of our approach comes from using the iterative conjugate gradient method to whiten the HOG template. Whitening a feature vector $x$ requires a system of linear equations $\Sigma w = (x - \mu)$ to be solved, usually by decomposing the covariance matrix $\Sigma$ (with LU or Cholesky decomposition). Since the covariance matrices of HOG templates have nice spectral properties (see Section \ref{sec:fast_whitening}), the conjugate gradient method will converge to the solution several orders of magnitude faster than decomposition methods.

@@

Table \ref{tab:who_initializations} provides a detailed comparison of WHO and NZ-WHO templates, illustrating that our method is faster and more discriminative than Whitened Histograms of Orientations. \scream{Does the table show this?}
% We compare the performance of WHO and NZ-WHO on 3D Object dataset \cite{Savarese07} and report performance on Table \ref{tab:who_initializations}.

@@

We use the GPU to parallelize the lookup of the correct autocovariance cells
during covariance matrix syntheses, which is \scream{x} times faster than using
the CPU (Figure \ref{fig:covariancetime}).

@@

Figure \ref{fig:spectral} shows the singular values of a covariance matrix for
one HOG template after a small regularization constant is added to the diagonal.
The condition number of of the matrix is \scream{x}, much smaller than the size
of the matrix (7500). As a result, conjugate gradient converges in \scream{x
ms}, two orders of magnitude faster than the system can be solved using the
Cholesky factorization (\scream{5 s}).

@@

\begin{comment}
\subsection{High Resolution Templates and FFT based Convolution}
\label{sec:fft} 
We generate high resolution templates with more than 250 HOG cells to capture details of an object to give accurate 2D-3D matching. These large templates cause computational burden when computing convolution. Though good for detecting accurate model and pose, convolution of high-resolution templates are much more slower since computation time scales linearly with the number of HOG cells in the template. To overcome and speed up the convolution, we propose FFT based GPU convolution which scales. Suppose there are length $n$ signal and length $m$ filter, naive convolution takes $O(nm)$ time where as FFT-based convolution takes $O\left( (n + m)\log (n+m) \right)$ time. For large $m$, high resolution template, we can gain computational advantage.

we pad zero to the 2D image features by the largest kernel size. And we transform the kernel to match the padded image feature. To compute convolution, we compute element-wise product with the transformed kernel and sum all the features. Unlike \cite{Podlozhnyuk}, we do not need to shift the data nor kernel if we use simple mathematical trick.

@@

[Figure for convolution time for different template resolution, CPU, GPU, naive convolution]
\end{comment}
%------------------------------------------------------------------------

\section{Estimating Parameters using MCMC}
\label{sec:fine}
% Estimating viewpoint continuously to fine tune a initial matching
The NZ-WHO template matching method we have presented makes template generation
and evaluation computationally inexpensive. This means we can efficiently
explore the parameter space to find the best object pose, scale, model type and
camera focal length.

Our method begins by convolving the

>>>>

We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), that are both
more discriminative than WHO, and can be generated in 70ms (compared to several
seconds for WHO). This speed up means we can generate templates on the fly,
allowing our approach to evaluate arbitrary viewpoint hypotheses.

\scream{chris:We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO),
an improvement over WHO which is both more discriminative than WHO and can be
computed many orders of magnitude faster than WHO (70ms as opposed to to several
seconds for WHO). This speedup allows us to generate NZ-WHO templates on the
fly, allowing us to quickly evaluate arbitrary viewpoint hypotheses.}

The computational efficiency and numerical stability of our approach comes from
using the iterative conjugate gradient method to whiten the HOG template.
Whitening a feature vector $x$ requires a system of linear equations $\Sigma w =
(x - \mu)$ to be solved, usually by decomposing the covariance matrix $\Sigma$
(with LU or Cholesky decomposition). Since the covariance matrices of HOG
templates have nice spectral properties (see Section \ref{sec:fast_whitening}),
the conjugate gradient method will converge to the solution several orders of
magnitude faster than decomposition methods.

@@

Table \ref{tab:who_initializations} provides a detailed comparison of WHO and
NZ-WHO templates, illustrating that our method is faster and more discriminative
than Whitened Histograms of Orientations. \scream{Does the table show this?}

% We compare the performance of WHO and NZ-WHO on 3D Object dataset \cite{Savarese07} and report performance on Table \ref{tab:who_initializations}.

@@

We use the GPU to parallelize the lookup of the correct autocovariance cells
during covariance matrix syntheses, which is an order of magnitude times faster than using
the CPU (Figure \ref{fig:covariancetime}).

@@

Figure \ref{fig:spectral} shows the singular values of a covariance matrix for
one HOG template after a small regularization constant is added to the diagonal.
The condition number of of the matrix is , much smaller than the size
of the matrix (7500). As a result, conjugate gradient converges in \scream{x
ms}, two orders of magnitude faster than the system can be solved using the
Cholesky factorization (\scream{5 s}).

@@

\subsection{High Resolution Templates and FFT based Convolution}
\label{sec:fft} 
We generate high resolution templates with more than 250 HOG cells to capture
details of an object to give accurate 2D-3D matching. Though good for detecting
accurate model and pose, convolution of high-resolution templates are much more
slower since computation time scales linearly with the number of HOG cells in
the template. To overcome and speed up the convolution, we propose FFT based
GPU convolution which scales. Suppose there are length $n$ signal and length
$m$ filter, naive convolution takes $O(nm)$ time where as FFT-based convolution
takes $O\left( (n + m)\log (n+m) \right)$ time. For large $m$, high resolution
template, we can gain computational advantage.

We pad zero to the 2D image features by the largest kernel size. And we
transform the kernel to match the padded image feature. To compute convolution,
we compute element-wise product with the transformed kernel and sum all the
features. Unlike \cite{Podlozhnyuk}, we do not need to shift the data nor
kernel if we use simple mathematical trick.

@@

% [Figure for convolution time for different template resolution, CPU, GPU, naive convolution]
%------------------------------------------------------------------------

\section{Continuous Viewpoint Estimation using MCMC}
\label{sec:fine}
% Estimating viewpoint continuously to fine tune a initial matching
The NZ-WHO template matching method we have presented makes template generation
and evaluation computationally inexpensive. This means we can efficiently
explore the parameter space to find the best object pose, scale, model type and
camera focal length.
